Black-hole ringdown as a probe of higher-curvature gravity theories

### 2022-04-13 papers/paper_alt_theor_bounds/paper_alt_theor_bounds.tex

% -----------
% Bibligraphy
% -----------
\AtBeginDocument{%
    \newwrite\bibnotes
    \def\bibnotesext{Notes.bib}
    \immediate\openout\bibnotes=\jobname\bibnotesext
    \immediate\write\bibnotes{@CONTROL{REVTEX41Control}}
    \immediate\write\bibnotes{@CONTROL{%
    apsrev41Control,author="08",editor="1",pages="1",title="0",year="1"}}
    \if@filesw
    \immediate\write\@auxout{\string\citation{apsrev41Control}}%
    \fi
}
% -----------

@@

% Comment commands
\usepackage{xcolor}
\newcommand{\agcomm}[1]{{\textcolor{red}{{[AG: #1]}} }}
\newcommand{\ag}[1]{{\textcolor{Maroon}{{#1}} }}
\newcommand{\hs}[1]{{\textcolor{blue}{{[HS: #1]}} }}
\newcommand{\ab}[1]{{\textcolor{green}{{[AB: #1]}} }}

@@

%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
% General:
The detection of gravitational waves produced by coalescing compact binaries
allows us to explore the dynamical, nonlinear regime of general relativity
and to constrain modifications to it.
%
Some of these gravitational-wave events are detected with sufficiently high
signal-to-noise ratio, allowing us to probe the relaxation of the remnant black
hole to its final, stationary state -- the so-called black-hole ringdown,
characterized by a set of quasinormal modes.
% Main question:
Can we use the ringdown to constrain deviations to general relativity,
as predicted by several of its contenders?
% What we do:
Here, we take a first step to address this question, by using an
inspiral-merger-ringdown waveform model in the effective-one-body formalism,
augmented with a parametrization of the ringdown based on an expansion on the
final black hole's spin.
%
We give a prescription on how to include to this waveform model the quasinormal mode frequencies
calculated on a theory-by-theory basis, in particular, for theories that deviate
from general relativity through higher-curvature corrections, namely,
Einstein-dilaton-Gauss-Bonnet, dynamical Chern-Simons, cubic- and quartic-order
effective-field-theories of general relativity.
%
% Main conclusion:
We confront the resultant waveform models against GW150914 and GW200129 data,
finding that while Einstein-dilaton-Gauss-Bonnet cannot be constrained with
these events, we can place upper bounds on the coupling constants of cubic-
($\ell_{\rm cEFT} \leqslant 38.2$~km) and quartic-order ($\ell_{\rm qEFT}
\leqslant 51.3$~km) effective-field-theories of general relativity, and
of dynamical Chern-Simons gravity ($\ell_{\rm CS} \leqslant 38.7$~km).
%
The latter result is a concrete example of a theory presently
unconstrained by inspiral-only analyses which, however, can be constrained by
merger-ringdown studies.
\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%

>>>>

% -----------
% Bibliography
% -----------
\AtBeginDocument{%
    \newwrite\bibnotes
    \def\bibnotesext{Notes.bib}
    \immediate\openout\bibnotes=\jobname\bibnotesext
    \immediate\write\bibnotes{@CONTROL{REVTEX41Control}}
    \immediate\write\bibnotes{@CONTROL{%
    apsrev41Control,author="08",editor="1",pages="1",title="0",year="1"}}
    \if@filesw
    \immediate\write\@auxout{\string\citation{apsrev41Control}}%
    \fi
}
% -----------

@@

% Comment commands
\usepackage{xcolor}
\newcommand{\agcomm}[1]{{\textcolor{red}{{[AG: #1]}} }}
\newcommand{\ag}[1]{{\textcolor{Maroon}{{#1}} }}
\newcommand{\hs}[1]{{\textcolor{blue}{{[HS: #1]}} }}
\newcommand{\ab}[1]{{\textcolor{cyan}{{#1}} }}
\newcommand{\abcomm}[1]{{\textcolor{red}{{[AB: #1]}} }}

@@

%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
% General:
The detection of gravitational waves produced by coalescing compact binaries
allows us to explore the dynamical, nonlinear regime of general relativity
and to constrain modifications to it.
%
Some of these gravitational-wave events are detected with sufficiently high
signal-to-noise ratio \ab{in the merger}, allowing us to probe the relaxation of the remnant black
hole to its final, stationary state --- the so-called black-hole ringdown,
characterized by a set of quasinormal modes.
% Main question:
Can we use the ringdown to constrain deviations to general relativity,
as predicted by several of its contenders?
% What we do:
Here, we \sout{take a first step to} address this question by using an
inspiral-merger-ringdown waveform model in the effective-one-body formalism,
augmented with a parametrization of the ringdown based on an expansion on the
final black hole's spin.
%
We give a prescription on how to include \ab{in} this waveform model the quasinormal mode frequencies
calculated on a theory-by-theory basis. \ab{In particular, we focus on gravity} theories that deviate
from general relativity through higher-order curvature corrections, namely,
Einstein-dilaton-Gauss-Bonnet, dynamical Chern-Simons, cubic- and quartic-order
effective-field-theories of general relativity.
%
% Main conclusion:
We confront the resultant waveform models against \ab{the two loudest gravitational-wave ringdown signals,
  notably, GW150914 and GW200129, and we find} that while the Einstein-dilaton-Gauss-Bonnet theory cannot be constrained with
these events, we can place upper bounds on the coupling constants of cubic-
($\ell_{\rm cEFT} \leqslant 38.2$~km) and quartic-order ($\ell_{\rm qEFT}
\leqslant 51.3$~km) effective-field-theories of general relativity, and
of dynamical Chern-Simons gravity ($\ell_{\rm CS} \leqslant 38.7$~km).
%
The latter result is a concrete example of a theory presently
unconstrained by inspiral-only analyses which, however, can be constrained by
merger-ringdown studies \ab{with current gravitational-wave data}.
\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%
